---
title: "CTD_timeseries"
author: "Matomo Niwano"
date: "3/23/2021"
output: html_document
---

## AMPLICON ANALYSIS (MiSeq Run: JBL8W)

```{console}

## PRIMER CLIPPING ##

# run cutadapt using custom script
# done on Ollie AWI server
# adjust for your own system

module load bio/cutadapt/3.2

bash ./../../../software/cutadapt.sh ./Original GTGYCAGCMGCCGCGGTAA CCGYCAATTYMTTTRAGTTT

## Create a text file with all sample names
ls -1 *R1_001.fastq.gz | sed 's/_R1_001\.fastq\.gz//' > ../sample_names.txt

```

## DADA2 processing
## Version of programs and R packages used in this analysis

- **R** = 4.0
- **R Studio** = 1.3.1093
- **dada2** = 1.14.1

```{r}
# Load packages
library(dada2)
library(ShortRead)
library(ggplot2)
library(gridExtra)
```

## Setting up the working directory

```{r}
# set working directory
setwd("/isibhv/projects/FRAMdata/MolObs/WaterCol_CTD_1519/JBL8W")

```

## Specify a path to the data
```{r}
# refinedData = primer-clipped fastq files
path <- "/isibhv/projects/FRAMdata/MolObs/WaterCol_CTD_1519/JBL8W/refinedData"
fns <- list.files(path)
fns
```

## Specifying file names for both forward and reverse files. 
```{r}
# ensure fwd/rev reads in same order
fnFs <- sort(list.files(path, pattern="R1_001.fastq.gz"))
fnRs <- sort(list.files(path, pattern="R2_001.fastq.gz"))

# Define sample names
sample.names <- sort(read.table("/isibhv/projects/FRAMdata/MolObs/WaterCol_CTD_1519/JBL8W/sample_names.txt", h = F, stringsAsFactors = F)$V1)

# Specify the full path to the fnFs and fnRs
fnFs <- file.path(path, fnFs)
fnRs <- file.path(path, fnRs)
```

## Quality check
```{r}
# Creating a pdf file that contains quality information for forward reads. 

QualityProfileFs <- list()
for(i in 1:length(fnFs)){QualityProfileFs[[i]] <- list()
  QualityProfileFs[[i]][[1]] <- plotQualityProfile(fnFs[i])}
pdf("JBL8W_QualityProfileForward.pdf")
for(i in 1:length(fnFs)) {
    do.call("grid.arrange", QualityProfileFs[[i]])}
dev.off()
rm(QualityProfileFs)

# Creating a pdf file that contains quality information for reverse reads
QualityProfileRs <- list()
for(i in 1:length(fnRs)) {
  QualityProfileRs[[i]] <- list()
  QualityProfileRs[[i]][[1]] <- plotQualityProfile(fnRs[i])}
pdf("JBL8W_QualityProfileReverse.pdf")
for(i in 1:length(fnRs)) {do.call("grid.arrange", 
                                  QualityProfileRs[[i]])}
dev.off()
rm(QualityProfileRs)

# Make directory and filenames for the filtered fastqs

filt_path <- file.path(path, "../Filtered")
if(!file_test("-d", filt_path)) dir.create(filt_path)
filtFs <- file.path(
  filt_path, paste0(sample.names, "_F_filt.fastq"))
filtRs <- file.path(
  filt_path, paste0(sample.names, "_R_filt.fastq"))

```

## Filtering and trimming
```{r}
# Filter: depending on expected overlap
#calculate expected error rates: #sum(10^(-x/10)) #x QC of each read
#truncLen lowered based on QualityProfile (low quality esp of rev-reads)

out <- filterAndTrim(
  fnFs,
  filtFs,
  fnRs,
  filtRs,
  truncLen = c(250, 195),
  maxN = 0,
  minQ = 2,
  maxEE = c(3, 3),
  truncQ = 0,
  rm.phix = T,
  compress = F,
  multithread = 6)
head(out)
summary(out[, 2]/out[, 1])



```
Min = 0.5697
1st Qu = 0.7983
Median = 0.8087
Mean = 0.8023
3rd Qu = 0.8163
Max = 0.8298

>80% (0.8087) This is good!


## Quality check after filtering and trimming

```{r}
# Forward files
QualityProfileFs.filt <- list()
for(i in 1:length(filtFs)){
  QualityProfileFs.filt[[i]] <- list()
  QualityProfileFs.filt[[i]][[1]] <- plotQualityProfile(
    filtFs[i])
  }
pdf("JBL8W_QualityProfileForwardFiltered.pdf")
for(i in 1:length(filtFs)) {
  do.call("grid.arrange", QualityProfileFs.filt[[i]])
}
dev.off()
rm(QualityProfileFs.filt)

# Reverse Files

QualityProfileRs.filt <- list()
for(i in 1:length(filtRs)) {
  QualityProfileRs.filt[[i]] <- list()
  QualityProfileRs.filt[[i]][[1]] <- plotQualityProfile(
    filtRs[i])
  }
pdf("JBL8W_QualityProfileReverseFiltered.pdf")
for(i in 1:length(filtRs)) {
  do.call("grid.arrange", 
    QualityProfileRs.filt[[i]])
  }
dev.off()
rm(QualityProfileRs.filt)

```
## Learn errors

```{r}
errF <- learnErrors(
  filtFs, multithread = 6, 
  randomize = T, verbose = 1, MAX_CONSIST = 20)
errR <- learnErrors(
  filtRs, multithread = 6, 
  randomize = T, verbose = 1, MAX_CONSIST = 20)

# Plot error profiles
pdf("JBL8W_ErrorProfiles.pdf")
plotErrors(errF, nominalQ = T)
plotErrors(errR, nominalQ = T)
dev.off()
# convergence should be after 5-8 rounds

# Dereplication 
derepFs <- derepFastq(filtFs, verbose = T)
derepRs <- derepFastq(filtRs, verbose = T)

# Name the derep-class objects by the sample names
names(derepFs) <- sample.names
names(derepRs) <- sample.names

# Denoising
dadaFs <- dada(
  derepFs, err = errF, multithread = 6, pool = "pseudo")
dadaRs <- dada(
  derepRs, err = errR, multithread = 6, pool = "pseudo")
```
95 samples were pooled: 12667640 reads in 1342240 unique sequences
95 sample were pooled: 12667640 reads in 1803800 unique sequences


## Read Merging

```{r}
# DADA recommends 20 for minOverlap parameter

mergers <- mergePairs(
  dadaFs,
  derepFs,
  dadaRs,
  derepRs,
  minOverlap = 20,
  verbose = TRUE,
  propagateCol = c(
    "birth_fold","birth_ham"))

# create a sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab) #95 10137

saveRDS(seqtab, "JBL8W.rds")

# save image
save.image("WaterCol_CTD_JBL8W.Rdata")

```

## Track reads through the pipeline

```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c(
"input","filtered","denoised","merged","tabled")
rownames(track) <- sample.names
track <- data.frame(track)

# Export the result to a text file
write.table(track, "dadastats_JBL8W.txt", quote=F, sep="\t")
```
